### 2024-08-15 QADT/notes.tex
Quotients of Algebraic Datatypes added

\documentclass[letterpaper,numbers=enddot]{scrartcl}

\usepackage{amsmath}
\usepackage{amssymb}

\title{Isomorphisms between algebraic datatypes}
\author{Andrew Polonsky, Ben Lenox}

\newcommand{\cat}{\mathbb{C}}
\newcommand{\op}{{op}}

\newcommand{\bone}{\boldsymbol{1}}
\newcommand{\then}{\quad\mathrel{\Longrightarrow}\quad}

\newcommand{\ints}{\mathbb{Z}}

\begin{document}

\maketitle

\section{2024.07.10}

Notes on ``Objects of categories as complex numbers''.

\subsection{Prop.2}

Can Proposition 2 be generalized as follows.

Let $p(x) = p_1(x)- p_2(x)$, $q(x) = q_1(x)-q_2(x)$ be elements of $\ints[x]$.

If the content of $p$ divides the content of $q$, and
the roots of $p$ are a submultiset of the roots of $q$, then
\[ \forall x. p(x)=0 \then q(x) = 0 \]

\newpage

\section*{TODO}

\begin{enumerate}
  \item Actually prove that
  \[ X=(X+1)^2 \then X \simeq X^2 \]
\end{enumerate}

\newpage
\section{Questions to be investigated}

\begin{enumerate}
  \item What is a strong isomorphism between ADTs?
  \item What does it mean for a functor to be a factor of another functor?
  \item A conception of a negative set in these terms.
  \begin{itemize}
    \item Universal $U$, and an ``integer set'' is a function from $U$ to $\mathbb{Z}$.
    \item A contravariant functor $- : \cat \to \cat^\op$ that splits the identity
    ($- \circ -^\op = id$)
    \item A solution to $E(X)=X$, obtained by taking $X = Y + 1$ for some type $Y$.
  \end{itemize}
  \item All polynomials functors with degree at least two will have linear factors.
  Can it solve arbitrary linear equation?
  Are all linear datatypes isomorphic to each other?  Are they generated by $S = -1$?
  \item Retractions between datatypes!
  \item Isomorphisms induced by initial algebra maps.
  \item Let $H(X) = 1 + X + X^3$, then $X$ contains sixth roots of unity as well as
  a square root of unity.  $M(X) = 1 + X + X^2$ contains fourth roots of unity;
  does this explain why we can ``divide $H(X)$ by $M(X)$''?

  Is this the reason why a solution to $M(X)=X$, when squared, yields a solution to $H(X)$?
  Why does this happen more generally?
  \item Can any equation $X=P(X)$, with $P(X)$ be an \emph{arbitrary}
  polynomial --- containing both positive and negative coefficients ---
  be solved by substitution $X = (Y+k)$, for some $k \ge 0$.

  Can various solutions be compared?

  \item Generalize all of the above to \emph{systems of equations} of the form
  \begin{align*}
    P_1(X_1,\dots,X_k) &= Q_1(X_1,\dots,X_k)\\
    P_2(X_1,\dots,X_k) &= Q_2(X_1,\dots,X_k)\\
     &\quad \vdots \\
    P_k(X_1,\dots,X_k) &= Q_k(X_1,\dots,X_k)
  \end{align*}
\end{enumerate}

\section{To do:}
\begin{enumerate}
  \item Define an ADT.
  \item Define R(A?)DT: \emph{Rational (Algebraic?) Datatypes}
  \item Is every RADT an ADT?
\end{enumerate}

\section{Discussion}
Let $F$ be a functor, $X$ be the initial $F$-algebra.

By Lambek's Theorem, $F(X) \simeq X$.

For any polynomial $P(X)$ in $X$, we get $P(F(X)) \simeq P(X)$.

Relate this to the notion of strong isomorphism.

\section{2023.08.10}

\subsection*{Functors}
Let $H(X) = 1 + X + X^3$.

Let $N$ be the initial algebra of $H$, so that $N \simeq H(N)$.

Let $B(X) = 1 + 2X$.

Let $S$ be the initial algebra of $B$, so that $S \simeq B(S)$.

Is $B$ a ``factor'' of $H$?

To compare the them, try to put an $B$-algebra on... $N$? Doesn't work.

But on $N^3$ works!

\section{2023.08.16}

\subsection{Main conjecture}

\subsection{Conjecture}

The algebraic datatype $X = P(X)$ includes infinite trees only if it does NOT contain
a positive integer root.

\textbf{Example.}
$F(X) = X^2-4X+1$.  To have $F(X)=X$, we expect the roots to be $X = 4$ and $X = 1$.

Applying the substitution $X = Y+1$, we get
\begin{align*}
  Y+1
  &= Y^2 + 2Y + 1 - 4Y-4 + 4\\
  Y &= Y^2-2Y
\end{align*}

The initial algebra for $Y \mapsto Y^2-2Y$ is $Y=\emptyset$,
which suggests the initial algebra for $F(X)$ is also $X = \bone$.

Setting instead $X = Z + 4$, we get
\[ Z+4 = \cdots = Z^2 + 4Z \]

Generally, whenever we set $X = Y + k$,
where $k$ is a root of the polynomial $F(X)$, then this will cancel out the constants,
leaving out a functor whose initial algebra is the empty set.

\subsection{Observation about $B(X) = 2X+1$}

\begin{align*}
  X &= 2X + 1\\
  X - 1 &= 2X\\
  X &= Y+1\\
  Y+1-1 &= 2Y + 2\\
  Y &= 2Y + 2
\end{align*}

\subsection{Key Question}

What is the notion of a \emph{least} fixed point/\emph{initial} algebra for
a (system of) equation(s) in which \emph{both sides} are polynomials.

(This is equivalent to admitting negative coefficients.)

\subsection{Another example}
\begin{align*}
  X &= X^2 - 4X + 4\\
  X &= 5 + P\\
  5+P &= \cdots \\
  P &= P^2 + 6P + 4
\end{align*}

\subsection{Open questions}

\begin{enumerate}
  \item The following rule is obviously sound:
  \[ X+1 = Y+1 \then X=Y \]
  How constructive is this principle?
  Do we need to decide equality on both $X$ and $Y$ to exhibit the
  isomorphism on the right in terms of the one on the left?

  \textbf{Answer.}
  Always solvable.  Given $i : X+1 \to Y+1$, define $i^! :X \to Y$ as follows.
  Given $x$, consider $X \mapsto X+1 \mapsto Y+1$, can check whether the output is in $1$, and if so,
  apply $i$ to $()$.  Same in the other direction.

  \item What does it mean for an isomorphism from $X$ to $P(X)$, where $P(X)$
  is a polynomial, to describe an ADT, when $P(X)$ has a specific root,
  e.g., $i$, $-i$, $1$, $-1$, $0$, $2$, $-\pi^2$, $\frac{1}{2}$, $\sqrt(2)$, etc.

  What properties does possession of a particular root imply for a given ADT?

  \item Generalize all of the above to having arbitrary ADTs as coefficients
  in the polynomial (defining the functor).

\end{enumerate}

\section{2023.09.07}
\begin{enumerate}
  \item If $X = P(X) + X$, then $X = n \cdot P(X) + X$ for all $n \in \mathbb{Z}$.

  \textsc{Proof.} Induction for non-negative $n$, symmetry for all $n$.

  \item If $X = P(X) + X$, then $X = X \cdot P(X) + X$.

  \item If $X = P(X) + X$, then $X = (X+n)\cdot P(X) + X$ for all $n \in \mathbb{Z}$.

  \textsc{Proof.} Distribute, and use the previous two facts.
\end{enumerate}

Everything reduces to the second case, so we analyze that.
\begin{itemize}
  \item If $P(x)=0$, it's trivially true.
  \item If $P(x) = k$, then $X = X + k$, $X \cdot P(X) + X = X (1 + k)$
\end{itemize}
\end{document}