### 2024-08-16 UTA/notes.tex
Added folder: UTA

\documentclass{scrartcl}

\title{Intersection Type Algebras}
\author{Aaron Frye, Andrew Polonsky}

\begin{document}

\section{2024.08.16}

A given set of equations can lead to multiple SR.  These are all equivalent.
\end{document}