### 2024-12-19 QADT/OpenProblems.tex

\documentclass{scrartcl}
\begin{document}

\begin{enumerate}
  \item{Commutativity of interpretations of EnvIso and Iso for ADTS}
  \item{A type theory defining equality without triple equality}
  \begin{itemize}
    \item{Isomorphisms without triple equality}
    \item{}
  \end{itemize}
  \item{}
\end{enumerate}

\end{document}