### 2025-01-22 UTA/UTA_Writeup.tex
Writeup added

\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\title{Formalizing Unification over Extensions of Simple Types}
\author{Aaron Frye}
\date{January 2025}

\begin{document}

\maketitle

\section{Abstract}
We use Agda to implement and verify type inference algorithms for extensions of simple types. We first treat recursive types and simple types, then subtyping.

\section{Formalization and Agda}
It is well known that the Curry-Howard correspondence draws an equivalence between proofs in constructive propositional logic and well typed programs with simple types. Agda leverages this fact to serve as both a functional programming language
and a way of writing formally verified mathematical proofs. It takes the idea a step further by incorporating dependent types, which allows the typing system to admit programs representing proofs in predicate logic.

Due to Agdas innate properties as a proof assistant, our results serve as both executable implementations of the listed algorithms and proofs of their correctness.

\section{Unification}
Unification in computer science refers to the problem of finding a substitution that solves a given set of equations. In our case we focus on unifying equations over types.

\section{Unification of recursive types}

\section{Rough ideas}
\begin{itemize}
    \item{unification}
    \item{Type Algebras}
    \item{Syntactic Algebras, notions of equivalence between algebras}
    \item{Free Algebras, invertibility}
    \item{Unification of Simple Types}
    \item{Deciding equivalence}
    \item{Recursive types}
    \item{The subtyping relation}
    \item{Intersection Types}
\end{itemize}

\end{document}