### 2025-02-11 Relations/Paper/Counterexamples.tex
New structure for the paper

\section{Counterexamples}
\label{sec:Counterexamples}