### 2025-02-11 Relations/Paper/Main.tex

\documentclass[letterpaper,USenglish,cleveref, autoref, thm-restate]{lipics-v2021}

\input{macros}

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Formalization of Abstract Rewriting}

\author{Andrew Polonsky}{Appalachian State University, USA}
{polonskya@appstate.edu} 
{} 
{}%TODO mandatory, 
% please use full name; only 1 author per \author macro; 
% first two parameters are mandatory, other parameters can be empty. 
% Please provide at least the name of the affiliation and the country. 
% The full address is optional. Use additional curly braces to indicate the correct
%  name splitting when the last name consists of multiple name parts.
\author{Samuel Arkle}{Appalachian State University, USA (Recent graduate)}
{arklesd@appstate.edu}{}{}

\Copyright{Andrew Polonsky and Samuel Arkle}

\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords

\supplement{https://github.com/DrPolonsky/LAM/tree/main/Relations}% TODO: set a repo just for this submissionoptional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...

\begin{document}
\maketitle

\begin{abstract}
    We present a constructive formalization of Abstract Reduction Systems (ARS)
     in the Agda proof assistant, focusing on the results given in Term Rewriting 
     Systems (Terese). We define a taxonomy of concepts related to termination and 
     confluence and investigate their relationships between each other and their 
     classical counterparts. We identify, and eliminate where possible, the use of classical 
     logic in the proofs of standard ARS results. Our analysis leads to refinements and mild
      generalizations of classical termination and confluence criteria.
\end{abstract}
\input{Introduction.tex}
\input{Definitions.tex}
\input{Implications.tex}
\input{Counterexamples.tex}
\input{Well-foundedness.tex}

\end{document}