### 2025-02-11 Relations/Paper/FinalTODO.tex
Added todo file.

\section{TODO}

I thought it would be useful to have a final todo in which we can list missing work or questions for this final report.

\begin{enumerate}
    \item Wording question. I'm trying to avoid stating "reduction" when talking about relations. However, can I still refer to a reduct and make sense? See for example the defintion of WCR.
    \item We have shown SN implies SMseq, can we show SN implies SM?
    \item Need to fill in ACM subject classifications
    \item Need to fill in keywords and phrases
    \item Need to check what other mandatory information is required in the file 
    \item Need to fill in Digital object identifier 
    \item Need to sort out a repository for the code and add it to the supplementary material section. 
    \item FINAL Remove this TODO section from the report 
\end{enumerate}

There will also be work we can do once we have submitted the file. We can jot that down here so we don't forget:

\begin{enumerate}
    \item Tidy up the code (fill in or remove goals. Remove unneeded comments. Make sure still compiles after moving it. Make sure it is consistent).
    \item We should create a good clear section of future work so we can pass this project on to future students.
\end{enumerate}