### 2025-02-21 Relations/Paper/Counterexamples.tex
Start on CE

\section{Counterexamples}
\label{sec:Counterexamples}

% Setting up the ce counter 
\newcounter{CEcounter}
\renewcommand{\theCEcounter}{\arabic{CEcounter}}

\newenvironment{counterexample}[1][]{%
    \refstepcounter{CEcounter} % Increment the counter
    \noindent \scriptsize\textbf{{Counterexample \theCEcounter.}} #1\par
}

This section collects the useful counterexamples we have encountered whilst undertaking this project. 
The counterexamples rule out certain implications between properties.

\begin{center}
\begin{tabular}{ccc}  % Three columns per row
    % First Counterexample
    \begin{minipage}{0.3\textwidth}
        \centering
        \begin{counterexample}\label{CE:1}
        \begin{tikzcd}[row sep=small, column sep=small]
            a & b \arrow[l] \arrow[r, bend left] & c \arrow[l, bend left] \arrow[r] & d
        \end{tikzcd} \\
    \end{counterexample}
    \end{minipage}
    &
    % Second Counterexample
    \begin{minipage}{0.3\textwidth}
        \centering
        \begin{counterexample}\label{CE:2}
        \begin{tikzcd}[row sep=small, column sep=small]
            a & b \arrow[l] \arrow[r, bend left] & c \arrow[l, bend left] \arrow[r] & d \arrow[loop right]
        \end{tikzcd} \\
        \end{counterexample}
    \end{minipage}
    &
    % Third Counterexample
    \begin{minipage}{0.3\textwidth}
        \centering
        \begin{counterexample}\label{CE:e}
        \begin{tikzcd}[row sep=small, column sep=small]
            a \arrow[loop left] & b \arrow[l] \arrow[r, bend left] & c \arrow[l, bend left] \arrow[r] & d \arrow[loop right]
        \end{tikzcd} \\
    \end{counterexample}
    \end{minipage}
    \\  % New row
    \\

    % Fourth Counterexample
    \begin{minipage}{0.3\textwidth}
        \centering
        \begin{counterexample}\label{CE:4}
        \begin{tikzcd}[row sep=small, column sep=small]
            n \in NF & a \arrow[l] \arrow[r] & b \arrow[r, bend left] & c \arrow[l, bend left]
        \end{tikzcd} \\
    \end{counterexample}
    \end{minipage}
    &
    % Fifth Counterexample
    \begin{minipage}{0.3\textwidth}
        \centering
        \begin{counterexample}\label{CE:5}
        \begin{tikzcd}[row sep=small, column sep=small]
            n \in NF & a \arrow[l, two heads] \arrow[r, two heads] & b \arrow[r, bend left] & c \arrow[l, bend left]
        \end{tikzcd} \\
    \end{counterexample}
    \end{minipage}
    &
    % Sixth Counterexample
    \begin{minipage}{0.3\textwidth}
        \centering
        \begin{counterexample}\label{CE:6}
        \begin{tikzcd}[row sep=small, column sep=small]
            a \arrow[r, bend left] & b \arrow[l, bend left]
        \end{tikzcd} \\
        \end{counterexample}
    \end{minipage}
    \\ % New row
    \\
    \begin{minipage}{0.3\textwidth}
        \centering
        \begin{counterexample}\label{CE:7}
        \begin{tikzcd}[row sep=small, column sep=small]
            f_0 \arrow[r] \arrow[dr] & f_1 \arrow[r] \arrow[d] & \dots \arrow[dl] \\
            & n \in NF 
        \end{tikzcd} \\
        \end{counterexample}
    \end{minipage}
    &
    \begin{minipage}{0.3\textwidth}
        \centering
        \begin{counterexample}\label{CE:8}
        \begin{tikzcd}[row sep=small, column sep=small]
            f_0 \arrow[r] \arrow[d] & f_1 \arrow[r] \arrow[d] & f_2 \arrow[r] \arrow[d] & \dots \arrow[d] \\
            n_0  & n_1  & n_2 & \dots 
        \end{tikzcd} \\
        \end{counterexample}
    \end{minipage}

    &
    \begin{minipage}{0.3\textwidth}
        \centering
        \begin{counterexample}\label{CE:9}
        \begin{tikzcd}[row sep=small, column sep=small]
            a \arrow[r, bend left] & b \arrow[l, bend left]
        \end{tikzcd} \\
        \end{counterexample}
    \end{minipage}

\end{tabular}
\end{center}