### 2025-02-23 Relations/Paper/Main.tex

\documentclass[letterpaper,USenglish,cleveref, autoref, thm-restate]{lipics-v2021}

\input{macros}

% \usepackage[dvipsnames]{xcolor}

\newcommand{\hired}[1]{{\color{red}{#1}}}

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Formalization of Abstract Rewriting}

\author{Samuel Arkle}{Appalachian State University, USA}
{arklesd@appstate.edu}{}{}
\author{Andrew Polonsky}{Appalachian State University, USA}
{polonskya@appstate.edu}
{}
{}%TODO mandatory,
% please use full name; only 1 author per \author macro;
% first two parameters are mandatory, other parameters can be empty.
% Please provide at least the name of the affiliation and the country.
% The full address is optional. Use additional curly braces to indicate the correct
%  name splitting when the last name consists of multiple name parts.

\authorrunning{S.\ Arkle and A.\ Polonsky}

\Copyright{Samuel Arkle and Andrew Polonsky}

\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords

\supplement{https://github.com/DrPolonsky/LAM/tree/main/Relations}% TODO: set a repo just for this submissionoptional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...

\begin{document}
\maketitle

\begin{abstract}
    We present a constructive formalization of Abstract Reduction Systems (ARS)
     in the Agda proof assistant, focusing on the results given in
     the standard reference on Term Rewriting Systems \cite{Terese}. We define a taxonomy of concepts related to termination and
     confluence and investigate their relationships between each other and their
     classical counterparts. We identify, and eliminate where possible, the use of classical
     logic in the proofs of standard ARS results. Our analysis leads to refinements and mild
      generalizations of classical termination and confluence criteria.
\end{abstract}

\input{Introduction.tex}
\input{Definitions.tex}
\input{Formalization.tex}
\input{Implications.tex}
\input{Well-foundedness.tex}
\input{Conclusion.tex}
\input{FinalTODO.tex}

\end{document}

Plan of the paper:

1. Introduction: 2 pages.
2. Definitions/setup: 2 pages. [WIP]
3. Formalization of Chapter 1 Terese, including Newman's Lemma: 2 pages [Still needs to be done]
- Statements (not proofs!) of key propositions and theorems in Agda code
4. Implications, hierachies, figures, and counterexamples: 4-5 [draft in report]
5. Classical principles and well-foundedness: 3 pages
6. Miscellaneous facts: 1 page
7. Related and future work: 1 page