### 2025-02-23 Relations/Paper/FinalTODO.tex

\section{TODO}

I thought it would be useful to have a final todo in which we can list missing work or questions for this final report.

\begin{enumerate}
    \item Need to fill in ACM subject classifications
    \item Establish [TeReSe] as how we refer to the book.
    \item Need to fill in keywords and phrases
    \item Need to check what other mandatory information is required in the file 
    \item Add bibliography/references.
    \item FINAL Remove this TODO section from the report 
\end{enumerate}

There will also be work we can do once we have submitted the file. We can jot that down here so we don't forget:

\begin{enumerate}
    \item Tidy up the code (fill in or remove goals. Remove unneeded comments. Make sure still compiles after moving it. Make sure it is consistent).
    \item We should create a good clear section of future work so we can pass this project on to future students.
    \item {We should remember that If global WCR and local SM implies local CR, then if we also have local WN, we get local SN.}
    \item \sacomment{Do we include more information on the propositions in the formalization section?}
    \item \sacomment{Proove the forawrd direction of theorem 1.2.3 iv}
    \item Show that SM and SMseq are classically equivalent.
\end{enumerate}