### 2025-02-28 Relations/Paper/FinalTODO.tex
updating ARS branch

\documentclass{scrartcl}

\input{packages.tex}
\input{macros.tex}

\begin{document}

\section{Feb 28 meeting}

\begin{enumerate}
  \item Add the implication RisWCR, x in WN and SM implies x in SN to Section 4.3.
  \item Clean up [TeReSe] references.
  \item Clean up the Well-foundedness
  \item Put up all classical principles in one place
  \item Organize the repo, hide private work.
  \item Clean up the Implications file, rename section etc.
  \item Getting rid of confluent/weakly confluent definition, standardize
\end{enumerate}

For the long term.
\begin{itemize}
  \item Knaster--Tarski, look over that
  \item Kleene tree to disprove isWFseq implies isWFacc
\end{itemize}

\section{TODO}

I thought it would be useful to have a final todo in which we can list missing work or questions for this final report.

\begin{enumerate}
    \item Need to fill in ACM subject classifications
    \item Establish [TeReSe] as how we refer to the book.
    \item Need to fill in keywords and phrases
    \item Need to check what other mandatory information is required in the file
    \item Add bibliography/references.
    \item FINAL Remove this TODO section from the report
\end{enumerate}

There will also be work we can do once we have submitted the file. We can jot that down here so we don't forget:

\begin{enumerate}
    \item Tidy up the code (fill in or remove goals. Remove unneeded comments. Make sure still compiles after moving it. Make sure it is consistent).
    \item We should create a good clear section of future work so we can pass this project on to future students.
    \item {We should remember that If global WCR and local SM implies local CR, then if we also have local WN, we get local SN.}
    \item \sacomment{Do we include more information on the propositions in the formalization section?}
    \item \sacomment{Proove the forawrd direction of theorem 1.2.3 iv}
    \item Show that SM and SMseq are classically equivalent.
\end{enumerate}

\end{document}