### 2025-08-06 Relations/Paper/WF_table.tex
Table improvements

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{array}
\usepackage{caption}
\usepackage{graphicx}  % <--- For \resizebox

\geometry{margin=1in}

\begin{document}

\begin{table}[h!]
\centering
\caption{Relational Table of Properties. Read along row and not down column. Note: Not determined whether being local or global makes an impact on the table.}
\renewcommand{\arraystretch}{1.3}
\resizebox{\textwidth}{!}{ % <--- This line scales the table
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
\hline
 & WFacc & WFseq & WFmin & WFminDNE& WFacc- & WFseq- & WFmin- & WFminDNE & WFminCor \\
\hline
WFacc     & - & Rdec & ? & ? & - & - & - & - & - \\
\hline
WFseq     & accCor + accDNE & - & ? & ? & accCor & - & accCor & accCor & - \\
\hline
WFmin     & {accDNE} & - & - & - & FB or accCor & - & accCor & accCor & - \\
\hline
WFminDNE  & {accDNE} & Rdec + MP & ? & - & accDNE or & - & - & - & - \\
\hline
WFacc-    & {accDNE} & accDNE + Rdec & ? & - & - & - & - & - & - \\
\hline
WFseq-    & accCor + {accDNE} & accCor + accDNE + Rdec & ? & ? & accCor & - & accCor & accCor & - \\
\hline
WFmin-    & (accCor or FB) + {accDNE} & (accCor or FB) + accDNE + Rdec & ? & ? & accCor or & - & - & - & - \\
\hline
WFminDNE-  & (accCor or FB) + {accDNE} & (accCor or FB) + accDNE + Rdec & ? & ? & accCor or & - & - & - & - \\
\hline
WFminCor  & ? & ? & ? & ? & ? & ? & ? & ? & - \\
\hline
\end{tabular}
} % <--- End resizebox
\end{table}

\end{document}