### 2025-11-05 Relations/Paper/Main.tex
Macro update, related works added to intro, main bibliography update.

\documentclass[letterpaper,USenglish,cleveref, autoref, thm-restate]{lipics-v2021}

\input{macros}

% \usepackage[dvipsnames]{xcolor}
\newtheorem{notation}[theorem]{Notation}

\newcommand{\hired}[1]{{\color{red}{#1}}}

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Formalization of Abstract Rewriting}

\author{Samuel Arkle}{Appalachian State University, USA}
{arklesd@appstate.edu}{}{}
\author{Andrew Polonsky}{Appalachian State University, USA}
{polonskya@appstate.edu}
{}
{}%TODO mandatory,
% please use full name; only 1 author per \author macro;
% first two parameters are mandatory, other parameters can be empty.
% Please provide at least the name of the affiliation and the country.
% The full address is optional. Use additional curly braces to indicate the correct
%  name splitting when the last name consists of multiple name parts.

\authorrunning{S.\ Arkle and A.\ Polonsky}

\Copyright{Samuel Arkle and Andrew Polonsky}

\ccsdesc[300]{Theory of computation~Type theory} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Abstract rewriting, Formalization, Confluence, Strong normalization, Newman's Lemma, Well-foundedness} %TODO mandatory; please add comma-separated list of keywords

\supplement{https://github.com/DrPolonsky/FPLA/tree/main}

\renewcommand\eminnershape{\bfseries}
\begin{document}
\maketitle

\begin{abstract}
    We present a constructive formalization of Abstract Reduction Systems (ARS)
     in the Agda proof assistant, focusing on the results given in
     the standard reference on Term Rewriting Systems \cite{Terese}. We define a taxonomy of concepts related to termination and
     confluence and investigate their relationships between each other and their
     classical counterparts. We identify, and eliminate where possible, the use of classical
     logic in the proofs of standard ARS results. Our analysis leads to refinements and mild
      generalizations of classical termination and confluence criteria.
      Finally, we investigate logical relationships between several notions of termination, 
      arising from different formulations of the concept of a well-founded relation.
    \end{abstract}

\input{Introduction.tex}
\input{Definitions.tex}
\input{Formalization.tex}
\input{Implications.tex}
\input{Well-foundedness.tex}
\input{Conclusion.tex}
% \input{FinalTODO.tex}

\begin{thebibliography}{10}
    \bibitem{Terese}
    Marc Bezem, Jan Willem Klop, and Roel de Vrijer.
    \newblock {Term Rewriting Systems}.
    \newblock {Cambridge University Press, 2003}.

    \bibitem{HoTT}
    Peter Aczel, Benedikt Ahrens, Thorsten Altenkirch, Steve Awodey,
    Bruno Barras, Andrej Bauer, Yves Bertot, Marc Bezem, Thierry Coquand,
    Eric Finster, et al.
    \newblock{Homotopy Type Theory: Univalent Foundations of Mathematics}.
    \newblock{The Univalent Foundations Program, Institute for Advanced Study, 2013}.

    \bibitem{QIIT}
    Thorsten Altenkirch, Paolo Capriotti, Gabe Dijkstra, Nicolai Kraus,
    and Fredrik Nordvall Forsberg.
    \newblock{Quotient inductive-inductive types}.
    \newblock{Proceedings of the International Conference on Foundations of Software Science and Computation Structures},
    \newblock{pages 293--310. Springer International Publishing, Cham, 2018}.

    \bibitem{CTT}
    Thierry Coquand, Simon Huber, and Anders Mörtberg.
    \newblock{On higher inductive types in cubical type theory}.
    \newblock{Proceedings of the 33rd Annual ACM/IEEE Symposium on Logic in Computer Science},
    \newblock{pages 255--264, 2018}.

    \bibitem{mimram}
    Mimram, Samuel
    \newblock{Program= Proof}.
    \newblock{2020}.
    \newblock{isbn = 9798615591839}.
    \newblock{Independently Published}.

    \bibitem{Bezem}
    Marc Bezem, Thierry Coquand, Keiko Nakata, and Erik Parmann.
    \newblock{Realizability at Work: Separating Two Constructive Notions of Finiteness}.
    \newblock {In 22nd International Conference on Types for Proofs and Programs (TYPES 2016)}.
    \newblock{Leibniz International Proceedings in Informatics (LIPIcs), Volume 97, pp. 6:1-6:23, Schloss Dagstuhl – Leibniz-Zentrum für Informatik (2018)}.
    \newblock{https://doi.org/10.4230/LIPIcs.TYPES.2016.6}.

    \bibitem{Berardi}
    Stefano Berardi
    \newblock{Classical and Intuitionistic Arithmetic with Higher Order Comprehension Coincide on Inductive Well-Foundedness}.
    \newblock{In 24th EACSL Annual Conference on Computer Science Logic (CSL 2015) }.
    \newblock{Leibniz International Proceedings in Informatics (LIPIcs), Volume 41, pp. 343-358}.
     \newblock{Schloss Dagstuhl – Leibniz-Zentrum für Informatik (2015)}.
      \newblock{https://doi.org/10.4230/LIPIcs.CSL.2015.343}.

    \bibitem{AgdaLibraryList}
    Agda Standard Library: List 
    \newblock{https://agda.github.io/agda-stdlib/v2.2/Agda.Builtin.List.html}
    \newblock{Accessed: 8th July 2025}

    \bibitem{AgdaLibraryRewriting}
    Agda Standard Library: Rewriting 
    \newblock{https://agda.github.io/agda-stdlib/v1.7.3/Relation.Binary.Rewriting.html}
    \newblock{Accessed: 5th November 2025}

    \bibitem{isabelleFormalization}
    Christian Sternagel and René Thiemann
    \newblock{https://www.isa-afp.org/entries/Abstract-Rewriting.html}
    \newblock{Accessed: 5th November 2025}

    \bibitem{trs}
    André Galdino, Mauricio Ayala-Rincón
    \newblock{http://trs.cic.unb.br/}
    \newblock{Accessed: 5th November 2025}

    \bibitem{coqARS}

    BLANQUI, FRÉDÉRIC, and ADAM KOPROWSKI. 
    \newblock{“CoLoR: A Coq Library on Well-Founded Rewrite Relations and Its Application to the Automated Verification of Termination Certificates.”}
    \newblock{Mathematical Structures in Computer Science 21, no. 4 (2011): 827–59.}
    \newblock{https://doi.org/10.1017/S0960129511000120.}

\end{thebibliography}

\end{document}

Plan of the paper:

1. Introduction: 2 pages.
2. Definitions/setup: 2 pages. [WIP]
3. Formalization of Chapter 1 Terese, including Newman's Lemma: 2 pages [Still needs to be done]
- Statements (not proofs!) of key propositions and theorems in Agda code
4. Implications, hierachies, figures, and counterexamples: 4-5 [draft in report]
5. Classical principles and well-foundedness: 3 pages
6. Miscellaneous facts: 1 page
7. Related and future work: 1 page